Estimating Learning Dynamics in the Routing Game

### 2015-10-12 paper/paper.tex

\documentclass{sig-alternate-ipsn13}
\usepackage{ntheorem}
\usepackage{algorithm}
\usepackage{algorithmic}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\input{latex_commands}
\newtheorem{remark}{Remark}
\newtheorem{theorem}{Theorem}
\newtheorem*{nonumtheorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}

@@

%============================================================================================
\newpage
\section{Experiment}
\label{sec:experiment}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{I}

Two versions of the game were presented to the user. One version of the game allowed each player to select the flow distribution for each path, the other version of the game only allowed the player to choose an ``exploration'' parameter, which selects the flow distribution for the players based on the KL-divergence.
The players interfaced with the game through a web interface where they can enter their inputs and receive feedback of each of their path cost and the cumulative cost.

The cost for each player's distribution were normalized by the calculated equilibrium flow to account for imbalances in the network.

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{II}
In the first version of the game, each player were allowed sliders to change how much of the flow are distributed to each path. All paths for a player have the same origin and destination node pair, but each player had unique origin and destination node pair.

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/no_slider_players_costs.png}
  \caption{Player normalized cost for no slider version.}
  \label{fig:cost_no_slider}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/no_slider_path_costs.png}
  \caption{Predicted path cost.}
  \label{fig:predicted_path_cost}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/no_slider_actual_flow_distribution.png}
  \caption{Player flow distribution.}
  \label{fig:no_slider_player_flow_distribution}
\end{figure}

We estimate the learning rate, $\hat \eta^k_t$,  of the players at turn $t$ by

\[
  \hat \eta^k_t = \arg\min_{\eta \geq 0} D_{\psi_k}(\hat x_k^{(t+1)}(\eta), x^{(t+1)}_k)
\]

From these estimated learning rate sequence, we then estimate the flow distribution of iteration $t+1$ using $\hat \eta^k_t $ and $x_k^{(t)}$ with \ref{eq:KL_update}.

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{III}

In the second version of the game, each player were only to change a single parameter $\eta$. This parameter controls how much each player want to explore from their previous flow distribution according to

\begin{equation} \label{eq:KL_update}
  (x_k^{(t+1)})_i = \frac{e^{-\eta_t^k (\ell^{(t)}_k)_i}}{\sum_j (x_k^{(t)})_i e^{-\eta_t^k (\ell^{(t)}_k)_i}}
\end{equation}

. The parameter for each turn is shown in \ref{fig:exploration}. We use a small $\epsilon$ to ensure that we will have no underflow for $x_k^{(t)}$ for numerical purposes. Then \ref{eq:KLupdate} becomes

\begin{equation} \label{eq:KL_update_epsilon}
  (x_k^{(t+1)})_i = \frac{e^{-\eta_t^k (\ell^{(t)}_k)_i}}{\sum_j (x_k^{(t)})_i e^{-\eta_t^k (\ell^{(t)}_k)_i}}
\end{equation}

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/slider_players_costs.png}
  \caption{Player normalized cost for slider version.}
  \label{fig:cost_slider}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/players_learning_rate.png}
  \caption{Player exploration rate.}
  \label{fig:exploration}
\end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=80mm]{images/slider_path_costs.png}
%   \caption{Player exploration rate.}
%   \label{fig:exploration}
% \end{figure}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------

>>>>

\documentclass{sig-alternate-ipsn13}
\usepackage{ntheorem}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{todonotes}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\input{latex_commands}
\newtheorem{remark}{Remark}
\newtheorem{theorem}{Theorem}
\newtheorem*{nonumtheorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}

% \usepackage{picins}
% \usepackage{graphicx}

% \makeatletter
% \def\Ginclude@graphics#1{%
%   \parpic(\Gin@@ewidth,\Gin@@eheight)[d]{#1}\picskip{0}}%
% \makeatother

@@

%============================================================================================
% \newpage
\section{Experiment}
\label{sec:experiment}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Game interface}

We simulate the proposed methods with a routing game interface that implements the behavior of the network. Players are assigned a unique origin and destination node pair and are tasked to assign a flow distribution to the paths between this pair at each turn of the game. The game calculates the cost along each edge of the network after all players have submitted their flow distribution for the turn. Each player then receives feedback with the local information of the cost for each path. The players then use this path cost history to assign a new flow distribution at the subsequent turns with the objective of minimizing their cumulative path cost.

\begin{figure*}
  \centering
  \includegraphics[width=160mm]{images/user_interface}
  \caption{Web interface for the routing game.}
  \label{fig:web_interface}
\end{figure*}

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\subsection{Convergence}

To address a player who may be at a disadvantage due to having longer paths or more congested paths because of the topology of the network, we normalized the cumulative path cost of each player with their cumulative path cost at equilibrium. We calculate this equilibrium path cost with Algorithm~\ref{alg:MD}.

At the beginning of the game, there is a clear exploration phase where players tend make aggressive assignments with their flow distribution. Towards the end of the game, the players consolidate their position and make more conservative flow distribution assignments, as seen with Figure~\ref{fig:normalized_costs}.

We can further show this consolidation from the players by computing the Rosenthal potential for the game. As the game progresses, we can see that the Rosenthal potential approaches the optimal value as seen in Figure~\ref{fig:global_potential}.

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/players_costs}
  \caption{Player normalized cost.}
  \label{fig:normalized_costs}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/global_potential_function}
  \caption{Rosenthal potential of the game.}
  \label{fig:global_potential}
\end{figure}

\subsection{Estimation and prediction}

We can estimate the learning rate of each player with the update in \ref{eq:estimation_eta}. We can see that the learning rates of some players are negative at certain turns, which may be explained by the players assuming a contrarian play style at flow distribution assignments for those certain turns.

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/learning_rate}
  \caption{Learning rate for each player.}
  \label{fig:learning_rates}
\end{figure}

With these estimated sequence of learning rates for each player, we can predict the flow distribution at turn $t+1$ with the learning rate $\eta_t$. We propose three methods for calculating $\eta_t$: first we let $\eta_t=\eta_{t-1}$, second we compute a moving average
\[
\eta_t=\dfrac{1}{N}\sum_{i=t-N-1}^{t-1}\eta_i
\] and finally we compute $\eta_t$ as a linear regression from $\eta_i,\ i \in [0,\ t-1]$.

We also compare these methods with a parameterized sequence of learning rate in Section 3.2

\begin{figure}
  \centering
  \includegraphics[width=80mm]{images/previous_eta_predictions}
  \caption{Flow distribution predictions $\eta_t = \eta_{t-1}$.}
  \label{fig:previous_eta_predictions}
\end{figure}

\begin{figure}
  \centering
  % \includegraphics[width=80mm]{images/fill_me}
  \missingfigure[figwidth=6cm]{Testing a long text string}
  \caption{Flow distribution predictions moving average $\eta_t$.}
  \label{fig:moving_eta_predictions}
\end{figure}

\begin{figure}
  \centering
  % \includegraphics[width=80mm]{images/fill_me}
  \missingfigure[figwidth=6cm]{Testing a long text string}
  \caption{Flow distribution predictions linear regression $\eta_t$.}
  \label{fig:linear_regression_eta_predictions}
\end{figure}

\begin{figure}
  \centering
  % \includegraphics[width=80mm]{images/fill_me}
  \missingfigure[figwidth=6cm]{Testing a long text string}
  \caption{Flow distribution predictions parameterized $\eta_t$.}
  \label{fig:parameterized_eta_predictions}
\end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=80mm]{images/no_slider_actual_flow_distribution.png}
%   \caption{Player flow distribution.}
%   \label{fig:no_slider_player_flow_distribution}
% \end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=80mm]{images/no_slider_KL_divergence_surface}
%   \caption{KL Divergence surface.}
%   \label{fig:kl_divergence_surface}
% \end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=80mm]{images/no_slider_global_potential_function}
%   \caption{Global potential function.}
%   \label{fig:global_potential_function}
% \end{figure}
